\documentclass{exam}

%------------------------ packages ------------------------%
\usepackage{amsmath,amsfonts,amsthm,amssymb,amsopn,bm}
\usepackage{pythontex}
\usepackage{fullpage}
\usepackage{graphicx}
\usepackage{fullpage}
\usepackage[paper=letterpaper,margin=1in,includeheadfoot,footskip=0.25in,headsep=0.25in]{geometry}
\usepackage{url}
\usepackage[usenames,dvipsnames]{color}
\usepackage[pdfborder={0 0 1},colorlinks=true,citecolor=black,plainpages=false]{hyperref}
%\usepackage{fancyhdr}
\usepackage{multirow}
\usepackage[english]{babel}
\usepackage{pdfpages,bbm}
\usepackage{enumitem}
\setlength {\marginparwidth }{2cm} 
\usepackage{todonotes}
\usepackage{tikz}
\usetikzlibrary{arrows,automata}
\usepackage[many]{tcolorbox}
\usepackage{array}
\usepackage{multirow}
\usepackage{listings}
\usepackage{color}
\usepackage{adjustbox}
\usepackage{wasysym} % For \CIRCLE
\renewcommand{\circle}{\tikz\draw[black] (0,0) circle (1ex);} % for \circle
\usepackage{cancel} % For \xcancel
\usepackage{forest}

%------------------------ math ------------------------%
\newcommand{\R}{\mathbb{R}} % real domain
\newcommand{\Rset}{\mathbb{R}} % real domain
\newcommand{\argmin}{\operatorname{argmin}}
\newcommand{\argmax}{\operatorname{argmax}}
\newcommand{\xv}{\mathbf{x}}
\newcommand{\av}{\mathbf{a}}
\newcommand{\bv}{\mathbf{b}}
\newcommand{\yv}{\mathbf{y}}
\newcommand{\wv}{\mathbf{w}}
\newcommand{\uv}{\mathbf{u}}
\newcommand{\zv}{\mathbf{z}}
\newcommand{\Xv}{\mathbf{X}}
\newcommand{\W}{\mathbf{W}}
\newcommand{\K}{\mathbf{K}}
\newcommand{\M}{\mathbf{M}}
\newcommand{\C}{\mathbf{C}}
\newcommand{\B}{\mathbf{B}}
\newcommand{\X}{\mathbf{X}}
\newcommand{\I}{\mathbf{I}}
\newcommand{\Proj}{\mathbf{P}}
\newcommand{\Y}{\mathbf{Y}}
\newcommand{\U}{\mathbf{U}}
\renewcommand{\L}{\mathbf{L}}
\newcommand{\y}{\mathbf{y}}
\newcommand{\x}{\mathbf{x}}
\newcommand{\w}{\mathbf{w}}
\newcommand{\mD}{\mathbf{D}}
\newcommand{\A}{\mathbf{A}}
\newcommand{\zero}{\mathbf{0}}
\newcommand{\norm}[1]{\left\lVert#1\right\rVert}
\DeclareMathOperator{\rank}{rank}
\newcommand{\E}{\mathbb{E}}


%------------------------ exam class macros ------------------------%
\checkboxchar{$\Box$}
% \renewcommand{\questionshook}{%
%     \setlength{\leftmargin}{10pt}%
%     \setlength{\labelwidth}{-\labelsep}%
% }
\renewcommand{\checkboxeshook}{
  \settowidth{\leftmargin}{W.}
  \labelwidth\leftmargin\advance\labelwidth-\labelsep
}

\newcommand{\grade}[1]{\small\textcolor{magenta}{\emph{[#1 points]}} \normalsize}
\newcommand{\sol}[1]{\ifshowsolutions{\leavevmode{\color{blue}Solution: #1}}\fi}

\newif\ifshowsolutions
% \showsolutionsfalse % whether to show solutions (true) or not (false)
\showsolutionstrue


\begin{document}

\title{Homework 4 - Part A}
\author{\Large \bf 10-405/10-605: Machine Learning with Large Datasets}
\date{{\bf Due Wednesday, March 31st at 11:59 PM Eastern Time}}
\maketitle

\paragraph{Instructions:} Submit your solutions via Gradescope, following the template below.  Note that this assignment is different from previous ones; it does not contain a theoretical written part and the programming part is much more open-ended than before. Because of this, we will not be autograding your submission. Instead, you will submit a report consisting of responses to questions asked in the notebook, along with your underlying code. The report (and the underlying code) are worth 100\% of your grade.

\paragraph{Submitting via Gradescope:} When submitting on Gradescope, you must assign pages to each question correctly (it
prompts you to do this after submitting your work). This significantly streamlines the
grading process for the course staff.  Failure to do this may result in a score of 0 for any questions
that you didn't correctly assign pages to. It is also your responsibility to make sure that your scan/submission is legible so that we can grade it.


\clearpage
\input{hw4/hw4_partA}

\clearpage
\input{collaboration}

\end{document}